---
title: "SCENIC"
author: "Özlem Bahar"
date: "6/9/2022"
output: html_document
---

```{r setup, include=FALSE, warning=FALSE}
knitr::opts_knit$set(root.dir = '/SCENIC_NBs')
knitr::opts_chunk$set(echo = TRUE)
if (!requireNamespace("BiocManager", quietly = TRUE)) install.packages("BiocManager")
BiocManager::version()
## Required
#BiocManager::install(c("AUCell", "RcisTarget"))
#BiocManager::install(c("GENIE3")) # Optional. Can be replaced by GRNBoost

require(AUCell)
#require(RcisTarget)
require(GENIE3)
require(zoo)
require(mixtools)
require(rbokeh)
require(DT)
require(NMF)
require(ComplexHeatmap)
require(R2HTML)
require(Rtsne)
require(doMC)
require(doRNG)
require(SCopeLoomR)
require(SCENIC)
require(Seurat)
require(miceadds)
require(SingleCellExperiment)
require(pheatmap)
```
## Download the species-specific databases for RcisTarget (the motif rankings). The links to all the available databases are available in https://resources.aertslab.org/cistarget/.

```{r}
 dbFiles <- c("https://resources.aertslab.org/cistarget/databases/drosophila_melanogaster/dm6/flybase_r6.02/mc8nr/gene_based/dm6-5kb-upstream-full-tx-11species.mc8nr.feather")
 dir.create("cisTarget_databases"); setwd("cisTarget_databases") # if needed
 for(featherURL in dbFiles)
 {   download.file(featherURL, destfile=basename(featherURL)) # saved in current dir
 }
# mc8nr: Motif collection version 8: 20k motifs
```

##Upload the species-specific databases for RcisTarget
```{r}
dbFiles <- importRankings("scenic_databases/dm6-5kb-upstream-full-tx-11species.mc8nr.feather")
```
##Formating the input

Retrieve the expression matrix and cell metadata (the function name depends on the object type or package).
```{r}
# Read the final R object and retrieve the expression matrix and cell metadata
#load .RDS object 
vnc_seurat<- readRDS(file="dpn.nb.clusters_list.rds")
DefaultAssay(vnc_seurat) <- 'RNA'
vnc_seurat <- NormalizeData(vnc_seurat)
exprMat <- vnc_seurat@assays$RNA@data
```

## create the cellInfo 
```{r}
cellInfo <- data.frame(CellType=Idents(vnc_seurat))
cellInfo$nGene <- colSums(exprMat>0)
cellInfo$nUMI <- vnc_seurat@meta.data$nUMI
cbind(table(cellInfo$CellType))
head(cellInfo)
```

```{r}
#dir.create("int")
cellInfo <- data.frame(cellInfo)
cbind(table(cellInfo$CellType))
saveRDS(cellInfo, file="int/cellInfo.Rds")
#cellInfo <- readRDS(file="int/cellInfo.Rds")
```

```{r, fig.height=5}
# Color to assign to the variables (same format as for NMF::aheatmap)
colVars <- list(CellType=c("7-4"="#F8766D",
                           "7-1"="#E58700",
                           "5-6"="#C99800", 
                           "5-3"="#A3A500", 
                           "5-2"="#6BB100",
                           "6-2"="#00BA38",
                           "4-2"="#00BF7D",
                           "3-5"="#00C0AF",
                           "2-5"="#00BCD8",
                           "3-2"="#00B0F6",
                           "1-2"="#619CFF",
                           "1-1"="#B983FF",
                           "MG"="#E76BF3",
                            "2-4"="#FD61D1",
                           "5-4"="#FF67A4"))

colVars$CellType <- colVars$CellType[intersect(names(colVars$CellType), cellInfo$CellType)]
#saveRDS(colVars, file="int/colVars.Rds")
plot.new(); legend(0,1, fill=colVars$CellType, legend=names(colVars$CellType))
```

```{r}
### Initialize settings
library(SCENIC)
org <- "dmel" 
dbDir <- "scenic_databases/" # RcisTarget databases location
myDatasetTitle <- "SCENIC_NBs" # choose a name for your analysis
data(defaultDbNames)
dbs <- defaultDbNames[[org]]
scenicOptions <- initializeScenic(org=org, dbDir=dbDir, dbs=dbs, datasetTitle=myDatasetTitle, nCores=10) 

# Databases:
scenicOptions@settings$dbs <- c("dm6-5kb-mc8nr"="dm6-5kb-upstream-full-tx-11species.mc8nr.feather")
#scenicOptions@settings$db_mcVersion <- "v8"
scenicOptions@inputDatasetInfo$cellInfo <- "int/cellInfo.Rds"
scenicOptions@inputDatasetInfo$colVars <- "int/colVars.Rds"

# Save to use at a later time...
saveRDS(scenicOptions, file="int/scenicOptions.Rds") 
```

```{r}
# Adjust the minimum according to the dataset
genesKept <- geneFiltering(as.matrix(exprMat), scenicOptions=scenicOptions,
                           minCountsPerGene=3*.01*ncol(exprMat),
                           minSamples=ncol(exprMat)*.01)
```

Check whether any known relevant genes are filtered-out (if any relevant gene is missing, double-check whether the filters are appropriate)

```{r}
interestingGenes <- c("en", "wg", "gsb", "hh", "Dr", "ind", "vnd")
# any missing? wg is missing from the raw data :( so that's expected
interestingGenes[which(!interestingGenes %in% genesKept)]
```
Filter the expression matrix to contain only these 5425 genes. This matrix is now ready for the co-expression analysis.

exprMat will not be used for the co-expression analysis.
```{r}
exprMat_filtered <- exprMat[genesKept, ]
dim(exprMat_filtered)
```
 
## Correlation matrix

GENIE3/GRNBoost can detect both positive and negative associations. In order to distinguish potential activation from repression, we will split the targets into positive- and negative-correlated targets (i.e. Spearman correlation between the TF and the potential target).

Calculate the correlation: 
```{r}
runCorrelation(as.matrix(exprMat_filtered), scenicOptions)
```
## Run GENIE3

Inputs include an expression matrix and a list of candidate regulators.
runGenie3 uses the TFs available in Rcistarget databases as candidate regulators.

```{r}
require(GENIE3)
# Run GENIE3
runGenie3(as.matrix(exprMat_filtered), scenicOptions)
```









